### 2014-05-07 docs/movement.tex
more writing

\documentclass{article}
\setlength{\parskip}{\baselineskip}%
\setlength{\parindent}{0pt}%
\begin{document}

The following are the movement vectors that have been defined in code.

\textbf{Cohesion}

Neighbors: team in sight by $r$ and $\alpha$ if they are not GUARDING or STUNNED.

\[ \vec { v_{ c } } =v_{ max }\frac { \Delta  \vec { p }  }{ \left\|   \Delta \vec { p }  \right\|  } \left( \frac { \left\|  \Delta \vec {  p }  \right\|  }{ r }  \right) ^{ 2 } \]

\textbf{Alignment}

Neighbors: team in sight by $r$ and $\alpha$ if they are SEEKING or SPREADING

\[ \vec { v_{ al } } =v_{ max }{\left( \frac { \left\| \Delta \vec { p } \right\|  }{ r }  \right)}^2\frac {\Delta \vec{v}} { \left\| \Delta \vec{v} \right\| } \]

\textbf{Avoidance}

Neighbors: enemies in sight by $r$ and $\alpha$ if not STUNNED

\[ \vec { v_{ av } } =\sum _{i=0}^{n} {v_{ max } \left( \frac{r - \left\| \Delta \vec{p_e} \right\|} {r} \right) \frac { \Delta \vec { p_e }  }{ \left\| \Delta \vec { p_e }  \right\|  }}  \]

\textbf{Separation}

Neighbors: team in sight by $r$ and $\alpha$ no matter their state

\[ \vec { v_{ d } } =-v_{ max }\frac {\Delta \vec{p}} { \left\| \Delta \vec{p} \right\| }{\left( \frac { r - \left\| \Delta \vec { p } \right\|  }{ r }  \right)}^2 \]

\textbf{Seeking/Homing/Mineral Cohesion}

Target: Some saved target either from memory or found in sight

\[ \vec{v_h} = v_{max} \frac {\vec{p_t} - \vec p} { \left\| \vec{p_t} - \vec p \right\| } \]

\textbf{Clearance}

Neighbors: team in sight by $r$ and $\alpha$ if they re not GUARDING or STUNNED

\[ \vec{v_{cl}} = v_{max} \frac {\Delta \vec p_\perp } { \left\| \Delta \vec p_\perp  \right\| }  \]

\end{document}